---
title: "Data used for this project"
output: html_notebook
---


Load data including only males. We are not authorized to make raw data publicly available.

```{r}
rm(list = ls())
FinPrisonMales <- 
  readRDS("C:/Users/benny_000/Dropbox/AAAKTUELLT/FinPrisonData/FinPrisonMales.rds")
```

We define a violent crime as crimes that belong to the homicide or assault categories. Other offence categories include offences that are violent and offences that are not violent. In many cases the violent forms are also coded as assault. 

We create variables for new violent crime

```{r}
FinPrisonMales <- 
  FinPrisonMales %>% 
  mutate(newO_violent = as.factor(ifelse(newO_homicide == 1 | newO_assault == 1, 
                                         "new_violent_crime", 
                                         "no_crime_or_not_violent")))

FinPrisonMales$newO_violent <- relevel(FinPrisonMales$newO_violent, 
                                       ref ="no_crime_or_not_violent")

table(FinPrisonMales$newO_violent)
```

Of the 748 new convictions 278 are categorized as violent.

Define predictor sets as character vectors 
