---
title: "Descriptive statistics"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package includes the data objects with descriptive statistics: `descriptive_stats_cat` (for categorical variables) and `descriptive_stats_num` (for numerical variables. These objects are data frames that can be used for further analyzes such as mean differences and chi-square tables.

For readers that only want to see the descriptive statistics we print out the tables below.


We can use the information in `variable_table` to replace the variable names with longer variable labels

```{r}
library(recidivismsl)

# Create a replacement vector
variable_labels <- variable_table$Label
names(variable_labels) <- variable_table$Variable
```


## Categorical variables
*Frequencies and percentages by reoffence group.Key: NoRe = No reoffence, NViolRe = Non-violent reoffence, VioRe = Violent Reoffence, n = frequencies, % = percentages.*

```{r}
printed_cat  <- descriptive_stats_cat
names(printed_cat) <- c("Variable", "Level", 
                         "NoRe n", "NViolRe n", "VioRe n",
                         "NoRe %", "NViolRe %", "VioRe %")



printed_cat$Variable <- variable_labels[printed_cat$Variable]


knitr::kable(printed_cat, digits = 1)

```

## Numerical variables
*Means and standard deviations per reoffence group. Key: NoRe = No reoffence, NViolRe = Non-violent reoffence, VioRe = Violent Reoffence, M = Means, SD = Standard Deviations.*

```{r}
printed_num  <- descriptive_stats_num
names(printed_num) <- c("Variable",
                         "NoRe M", "NViolRe M", "VioRe M",
                         "NoRr SD", "NViolRe SD", "VioRe SD")



printed_num$Variable <- variable_labels[printed_num$Variable]


knitr::kable(printed_num, digits = 1)

```



