---
title: "Incremental validity in Cohen's d"
author: "Benny Salo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette we show how we used the results from the 1000 resamples in the training set to compare the overall performance of a two competing models. The comparisons are of two types:
   - for the models with the same predictor set but using different algorithms (elastic net vs. random forest)
   - for models using elastic net between models with an incremetally larger set of predictors. (These results are presented in Table 2.)
   
We use McFadden's pseudo R-squared as a measure of overall performance but transform these values to d before vomparing models.
   

```{r}
library(recidivismsl)
library(dplyr)
```

## Function

We will use this function to compare an incremental model to a simpler model. It will:
   - extract the columns in the `values`element of `model_perfs_training_set1000` that refer to the McFadden pseudo R2 values for the incremental (`incr_mod`) and the simpler models (`simpler_mod`). (It will do this via using column names that are specified later.)
   - transform these R2 values to d. (See `?calc_d_from_R2()`)
   - calculate the median, percetiles 2.5 and 97.5, and the proportion of cases when the difference is 0 or negative (interpreted as a one-sided p_value) 

```{r}
R2_difference_in_d <- function(incr_mod, simpler_mod, 
                               resamples = model_perfs_training_set1000) {
  
  data_frame(R2_b = resamples$values[[incr_mod]],
             R2_w = resamples$values[[simpler_mod]]) %>% 
    mutate(incr_d = calc_d_from_R2(R2_b),
           simpler_d  = calc_d_from_R2(R2_w),
           diff_d     = incr_d - simpler_d) %>%
    summarise(median_diff_d = median(diff_d),
              LL            = quantile(diff_d, 0.025),
              UL            = quantile(diff_d, 0.975),
              p_one_sided   = mean(diff_d <= 0))
}
```

## Comparing algorithm types

Create a data frame with pairs of models to compare: corresponding elastic net (glmnet) and random forest (rf) models.

```{r}
compared_models <- 
  model_grid %>% 
  filter(analysis == "Main analyses" & model_type != "Logistic regression") %>% 
  select(model_name) %>% 
  purrr::as_vector()


model_pairs_algo <- 
  data_frame(glmnet_mod = stringr::str_subset(compared_models, "glmnet"),
             rf_mod     = stringr::str_subset(compared_models, "rf")) %>%
  mutate(glmnet_mod = paste0(glmnet_mod, "~McF_R2"),
         rf_mod     = paste0(rf_mod, "~McF_R2")) 


```

Make the comparisons.

```{r}
purrr::map2_df(.x = model_pairs_algo$glmnet_mod, 
               .y = model_pairs_algo$rf_mod, 
               .f = ~R2_difference_in_d(.x, .y)) %>% 
  # Bind in columns with compared models
  bind_cols(model_pairs_algo, .) %>% 
  # Clean up model names
  mutate_at(c("glmnet_mod", "rf_mod"), ~stringr::str_remove(., "~McF_R2"))
```

## Comparing predictor sets

Create a data frame with pairs of models to compare: models with same outcome and incremental predictor sets.

```{r}
incremental_mod <- c("gen_stat_glmnet",
                     "gen_bgnn_glmnet",
                     "gen_allp_glmnet",
                     
                     "vio_stat_glmnet",
                     "vio_bgnn_glmnet",
                     "vio_allp_glmnet")

baseline_mod    <- c("gen_rita_glmnet",
                     "gen_stat_glmnet",
                     "gen_bgnn_glmnet",
                     
                     "vio_rita_glmnet",
                     "vio_stat_glmnet",
                     "vio_bgnn_glmnet")




```

Make the comparisons.

```{r}
model_pairs_preds <- 
  data_frame(incr = incremental_mod,
             base = baseline_mod) %>% 
      mutate(incr = paste0(incr, "~McF_R2"),
             base = paste0(base, "~McF_R2")) 

purrr::map2_df(.x = model_pairs_preds$incr, 
               .y = model_pairs_preds$base, 
               .f = ~R2_difference_in_d(.x, .y)) %>% 
  bind_cols(model_pairs_preds, .) %>% 
  mutate_at(c("incr", "base"), ~stringr::str_remove(., "~McF_R2"))
```
