---
title: "Table 2 - Pseudo R2"
date: "`r Sys.Date()`"
output: github_document
---

In this vignette we show how we extracted McFadden pseudo R2-values for Table 2. 

Model performances in the training set are stored in the data object `model_perfs_training_set1000`. It also contains other models and measures than the ones presented in Table 2. (See the documentation for the data object.)

```{r}
?model_perfs_training_set1000
```

Load packages.

```{r}
library(recidivismsl)
library(dplyr)
```

We will extract the columns with McFadden pseudo R2 values from the 'values' element in the data object. After that we will tidy the data frame into long format and calculate for each model the median the 2.5 and 97.5 percentiles plus the proportion of values that are under 0 (interpreted as a one-sided p-value).



```{r}
# Functions from dplyr used (select, mutate, group_by, summarise)
McF_tbl_cv <-
model_perfs_training_set1000$values %>% 
  # Extract columns (one for each model)
  select(Resample, ends_with(match = "McF_R2")) %>% 
  # Tidy into long format
  tidyr::gather(-Resample, key = model, value = McF_R2) %>%
  # Calculate median, percentiles and one-sided p
  group_by(model) %>% 
  summarise(median_McF_R2 = median(McF_R2),
            ci95_LL_R2    = quantile(McF_R2, probs = 0.025),
            ci95_UL_R2    = quantile(McF_R2, probs = 0.975),
            p_one_side    = mean(McF_R2 < 0)) %>% 
  # Remove ~ROC from the end of the model name
  mutate(model = stringr::str_replace(model, pattern = "~McF_R2", ""))


```

We can clean up this table a bit for printing.

```{r}
# Get som model descriptions from `model_grid`
model_desc <- model_grid[c("model_name", "outcome", 
                           "predictors", "model_type", "analysis")]


table_2 <- 
  # Add the descriptions
  full_join(model_desc, McF_tbl_cv, by = c("model_name" = "model")) %>% 
  # Only keep Elastic net models and the models using individual dimensions.
  filter(model_type == "Elastic net" | analysis == "Dimension analyses") %>% 
  # Sort the table
  arrange(outcome, desc(analysis), predictors) %>% 
  # Remove superfluous columns
  select(-model_name, -analysis, -model_type) %>%
  # Round to three digits
  mutate_at(.vars = vars(3:5), round, 3)
```

Print the table.

```{r}
knitr::kable(table_2)
```








